Denotational Semantics for SSA with Weak Memory Operations

### 2023-10-04 pldi24/paper.tex
Began copying over typing rules
Copied over typing rules

% Syntax
\newcommand{\letexpr}[3]{\ensuremath{\ms{let}\;#1 = #2\;\ms{in}\;#3}}
\newcommand{\letstmt}[3]{\ensuremath{\ms{let}\;#1 = #2; #3}}
\newcommand{\jmpb}[2]{\ms{jmp}\;#1\;#2}
\newcommand{\ljmpb}[2]{\jmpb{\lbl{#1}}{#2}}
\newcommand{\ite}[3]{\ms{if}\;#1\;\{#2\}\;\ms{else}\;\{#3\}}
\newcommand{\ewhere}[2]{\ms{then}\;#1\;\ms{where}\;#2}
\newcommand{\where}[2]{#1\;\ms{where}\;#2}
\newcommand{\wbranch}[3]{#1(#2) \lto #3}
\newcommand{\lwbranch}[3]{\wbranch{\lbl{#1}}{#2}{#3}}
\newcommand{\bsplice}[3]{#1(#2)\;\{#3\}}
\newcommand{\lbsplice}[3]{\bsplice{\lbl{#1}}{#2}{#3}}

@@

\begin{figure}
  \begin{center}
    \begin{grammar}
      <\(a, b, c, e\)> ::= \(x\) 
      \;|\; \(f\;a\)
      \;|\; \((a, b)\) 
      \;|\; \(()\) 
      \;|\; \(\ctt\) 
      \;|\; \(\cff\)
      \;|\; \(\letexpr{x}{a}{e}\)
      \;|\; \(\letexpr{(x, y)}{a}{e}\)
      \;|\; \(\lbsplice{\ell}{x: A}{t}\)

      <\(s, t\)> ::= \(\ljmpb{\ell}{a}\) 
      \;|\; \(\ite{e}{s}{t}\)
      \;|\; \(\letstmt{x}{a}{t}\)
      \;|\; \(\letstmt{(x, y)}{a}{t}\)
      \;|\; \(\ewhere{t}{L}\)

@@

      <\(\Gamma\)> ::= \(\cdot\) \;|\; \(\Gamma, x: A\)

@@

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\Gamma \vdash_p a: A\) &
        \(a\) is a term of type \(A\) in context \(\Gamma\) with purity \(p \in \{0, 1\}\) \\
        \(\Gamma \vdash t \rhd \ms{L}\) &
        \(t\) is a block targeting label-set \(\ms{L}\) in context \(\Gamma\) \\
        \(\ms{L} \vdash L \rhd \ms{K}\) &
        The labels \(L\) send label-set \(\ms{L}\) to label-set \(\ms{K}\)
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa}
  \Description{Typing judgements for isotope-SSA}
  \label{fig:ssa-judgements}
\end{figure}

\begin{figure}
  \begin{center}    
    \TODO{this}
  \end{center}
  \caption{Typing rules for \isotopessa}
  \Description{Typing rules for isotope-SSA}
  \label{fig:ssa-typing}
\end{figure}

>>>>

% Syntax
\newcommand{\letexpr}[3]{\ensuremath{\ms{let}\;#1 = #2\;\ms{in}\;#3}}
\newcommand{\letstmt}[3]{\ensuremath{\ms{let}\;#1 = #2; #3}}
\newcommand{\brb}[2]{\ms{br}\;#1\;#2}
\newcommand{\lbrb}[2]{\brb{\lbl{#1}}{#2}}
\newcommand{\ite}[3]{\ms{if}\;#1\;\{#2\}\;\ms{else}\;\{#3\}}
\newcommand{\ewhere}[2]{\ms{then}\;#1\;\ms{where}\;#2}
\newcommand{\where}[2]{#1\;\ms{where}\;#2}
\newcommand{\wbranch}[3]{#1(#2) \lto #3}
\newcommand{\lwbranch}[3]{\wbranch{\lbl{#1}}{#2}{#3}}
\newcommand{\bsplice}[3]{#1(#2)\;\{#3\}}
\newcommand{\lbsplice}[3]{\bsplice{\lbl{#1}}{#2}{#3}}
\newcommand{\csplits}[3]{#1 \mapsto #2;#3}
\newcommand{\cwk}[2]{#1 \mapsto #2}
\newcommand{\lwk}[2]{#1 \rightsquigarrow #2}
\newcommand{\tlin}[2]{\ms{lin}_{#2}(#1)}
\newcommand{\ltlin}[3]{\ms{lin}_{#3}(#1^{#2})}
\newcommand{\thyp}[3]{#1: {#2}^{#3}}
\newcommand{\lhyp}[3]{#1[#2](#3)}
\newcommand{\llhyp}[3]{\lhyp{\lbl{#1}}{#2}{#3}}
\newcommand{\rle}[1]{{\scriptsize\textsf{#1}}}
\newcommand{\taff}{\ms{a}}
\newcommand{\trel}{\ms{r}}
\newcommand{\tint}{\infty}
\newcommand{\hasty}[4]{#1 \vdash_{#2} #3: #4}
\newcommand{\haslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\lhaslb}[3]{#1 \vdash #2 \rhd #3}

@@

\begin{figure}
  \begin{center}
    \begin{grammar}
      <\(A, B, C\)> ::= \(X\)
      \;|\; \(A \otimes B\)

      <\(a, b, c, e\)> ::= \(x\) 
      \;|\; \(f\;a\)
      \;|\; \((a, b)\) 
      \;|\; \(()\) 
      \;|\; \(\ctt\) 
      \;|\; \(\cff\)
      \;|\; \(\letexpr{x}{a}{e}\)
      \;|\; \(\letexpr{(x, y)}{a}{e}\)
      \;|\; \(\lbsplice{\ell}{x: A}{t}\)

      <\(s, t\)> ::= \(\lbrb{\ell}{a}\) 
      \;|\; \(\ite{e}{s}{t}\)
      \;|\; \(\letstmt{x}{a}{t}\)
      \;|\; \(\letstmt{(x, y)}{a}{t}\)
      \;|\; \(\ewhere{t}{L}\)

@@

      <\(\Gamma\)> ::= \(\cdot\) \;|\; \(\Gamma, \thyp{x}{A}{q}\)

@@

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\hasty{\Gamma}{p}{x}{A}\) &
        \(a\) is a term of type \(A\) in context \(\Gamma\) with purity \(p \in \{0, 1\}\) \\
        \(\haslb{\Gamma}{t}{\ms{L}}\) &
        \(t\) is a block targeting label-set \(\ms{L}\) in context \(\Gamma\) \\
        \(\lhaslb{\ms{L}}{L}{\ms{K}}\) &
        The labels \(L\) send label-set \(\ms{L}\) to label-set \(\ms{K}\) \\
        \(\csplits{\Gamma}{\Delta}{\Xi}\) &
        The context \(\Gamma\) splits into \(\Delta\) and \(\Xi\) \\
        \(\tlin{A}{q}\) &
        The type \(A\) has linearity \(q\) \\
        \(\ltlin{A}{r}{q}\) &
        \(\tlin{A}{q} \land q \subseteq r\) \\
        \(\tlin{\Gamma}{q}\) &
        The context \(\Gamma\) has linearity \(q\) \\
        \(\cwk{\Gamma}{\Delta}\) &
        \(\Gamma\) is a weakening of \(\Delta\) (i.e. \(\csplits{\Gamma}{\Delta}{\cdot}\)) \\
        \(\lwk{\ms{L}}{\ms{K}}\) &
        The label-set \(\ms{L}\) weakens to the label-set \(\ms{K}\)
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Typing judgements for \isotopessa}
  \Description{Typing judgements for isotope-SSA}
  \label{fig:ssa-judgements}
\end{figure}

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{base-lin}}{q \subseteq \ms{lin}(X)}{\tlin{X}{q}} \qquad
    \prftree[r]{\rle{pair-lin}}{\tlin{A}{q}}{\tlin{B}{q}}{\tlin{A \otimes B}{q}} \qquad
    \prftree[r]{\rle{nil-lin}}{\tlin{\cdot}{q}} \qquad
    \prftree[r]{\rle{cons-lin}}{\ltlin{A}{r}{q}}{\tlin{\Gamma}{q}}
      {\tlin{\Gamma, \thyp{x}{A}{r}}{q}} \\
    \prftree[r]{\rle{split-nil}}{\csplits{\cdot}{\cdot}{\cdot}} \qquad
    \prftree[r]{\rle{split-left}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {r \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta, \thyp{x}{A}{r}}{\Xi}} \qquad
    \prftree[r]{\rle{split-right}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {r \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta}{\Xi, \thyp{x}{A}{r}}} \\
    \prftree[r]{\rle{split-dup}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\ltlin{A}{q}{\trel}}
      {r, s \subseteq q}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta, \thyp{x}{A}{r}}{\Xi, \thyp{x}{A}{s}}}
      \qquad
    \prftree[r]{\rle{split-drop}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\ltlin{A}{q}{\taff}}
      {\csplits{\Gamma, \thyp{x}{A}{q}}{\Delta}{\Xi}}
      \\
    \prftree[r]{\rle{join-nil}}{\lwk{\cdot}{\cdot}} \qquad
    \prftree[r]{\rle{join-cons}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}} 
      \qquad
    \prftree[r]{\rle{join-zero}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\lwk{\ms{L}}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}} 
  \end{gather*}
  \caption{Structural rules for \isotopessa}
  \Description{Structural rules for isotope-SSA}
  \label{fig:ssa-structural}
\end{figure}

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{var}}
      {\cwk{\Gamma}{\thyp{x}{A}{q}}}
      {\hasty{\Gamma}{p}{x}{A}} \qquad
    \prftree[r]{\rle{app}}
      {f \in \mc{I}_p(A, B)}
      {\hasty{\Gamma}{1}{a}{A}}
      {\hasty{\Gamma}{p}{f\;a}{B}} \qquad
    \prftree[r]{\rle{pair}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta}{1}{a}{A}}
      {\hasty{\Xi}{1}{b}{B}}
      {\hasty{\Gamma}{1}{(a, b)}{A \otimes B}} \\
    \prftree[r]{\rle{unit}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{1}{()}{\mb{1}}} \qquad
    \prftree[r]{\rle{true}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{1}{\ctt}{\mb{2}}} \qquad
    \prftree[r]{\rle{false}}
      {\cwk{\Gamma}{\cdot}}
      {\hasty{\Gamma}{1}{\cff}{\mb{2}}} \\
    \prftree[r]{\rle{let}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta, \thyp{x}{A}{}}{p}{e}{B}}
      {\hasty{\Xi}{1}{a}{A}}
      {\hasty{\Gamma}{p}{\letexpr{x}{a}{e}}{B}} \qquad
    \prftree[r]{\rle{blk}}
      {\haslb{\Gamma}{t}{\llhyp{\ell}{\cdot}{A}}}
      {\hasty{\Gamma}{0}{\lbsplice{\ell}{A}{t}}{A}} \\
    \prftree[r]{\rle{let2}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{p}{e}{C}}
      {\hasty{\Xi}{1}{a}{A \otimes B}}
      {\hasty{\Gamma}{p}{\letexpr{(x, y)}{a}{e}}{C}}
  \end{gather*}
  \caption{Typing rules for \isotopessa terms}
  \Description{Typing rules for isotope-SSA terms}
  \label{fig:ssa-term-typing}
\end{figure}

\begin{figure}
  \begin{gather*}    
    \prftree[r]{\rle{br}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\lwk{\llhyp{\ell}{\Delta}{A}}{\ms{L}}}
      %TODO: consider allowing binding here, or should it be purely for lets?
      {\hasty{\Xi}{1}{a}{A}}
      {\haslb{\Gamma}{\lbrb{\ell}{a}}{\ms{L}}} \qquad
    \prftree[r]{\rle{ite}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\hasty{\Delta}{1}{e}{\mb{2}}}
      {\haslb{\Xi}{s}{\ms{L}}}
      {\haslb{\Xi}{t}{\ms{L}}}
      {\haslb{\Gamma}{\ite{e}{s}{t}}{\ms{L}}} \\
    \prftree[r]{\rle{let-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\haslb{\Delta, \thyp{x}{A}{}}{t}{\ms{L}}}
      {\hasty{\Xi}{p}{a}{A}}
      {\haslb{\Gamma}{\letstmt{x}{a}{t}}{\ms{L}}} \\
    \prftree[r]{\rle{let2-blk}}
      {\csplits{\Gamma}{\Delta}{\Xi}}
      {\haslb{\Delta, \thyp{x}{A}{}, \thyp{y}{B}{}}{t}{\ms{L}}}
      {\hasty{\Xi}{p}{a}{A \otimes B}}
      {\haslb{\Gamma}{\letstmt{(x, y)}{a}{t}}{\ms{L}}} \\
    \prftree[r]{\rle{where}}
      {\haslb{\Gamma}{t}{\ms{L}}}
      {\lhaslb{\ms{L}}{L}{\ms{K}}}
      {\haslb{\Gamma}{\ewhere{t}{L}}{\ms{K}}} \qquad
    \prftree[r]{\rle{nil-br}}
      {\lwk{\ms{L}}{\ms{K}}}
      {\lhaslb{\ms{L}}{\cdot}{\ms{K}}} \qquad
    \prftree[r]{\rle{cons-br}}
      {\lhaslb{\ms{L}}{L}{\ms{K}, \llhyp{\ell}{\Gamma}{A}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{t}{\ms{L}}}
      {\lhaslb{\ms{L}}{L, \lwbranch{\ell}{x: A}{t}}{\ms{K}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa blocks}
  \Description{Typing rules for isotope-SSA blocks}
  \label{fig:ssa-block-typing}
\end{figure}
